---
title: "HW04_Sampathirao_Anvita"
author: "Anvita Sampathirao"
date: "6/4/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
#1a.
$$z=\frac{x-\mu}{\sigma}$$
```{r}
z <- (45-70)/10
z
```
#1b.
```{r}
pnorm(45,mean = 70, sd =10, lower.tail = TRUE)
```
#1c.
```{r}
2*pnorm(45,mean = 70, sd =10, lower.tail = TRUE)
```
#2a.
```{r}
SMPL<- as.data.frame(matrix(1:100, nrow = 10000, ncol = 10))
RNDM<- function(d, x) {
    s <- c(d[sample(nrow(d), 1), sample(ncol(d), 1)])
    x <- x - 1  
    for (i in c(1:x)) {
        row <- sample(nrow(d), 1)  
        col <- sample(ncol(d), 1) 
        s<- append(s, d[row, col], after = length(s))
        i <- i + 1  
    }
    return(s)
}
s1 <- RNDM(SMPL, 10)
s1
```
#2b.
$$\bar{x}=\frac{1}{N}\sum_{i=1}^{N}{x_i}$$
```{r}
mean1 <- sum(s1)/length(s1)
mean1
```
#2c.
$$s=\sqrt{\sum_{i=1}^{n}\frac{(x_i-\bar{x})^2}{n-1}}$$
```{r}
standarddev<- sqrt(sum((s1-mean1)^2/(length(s1)-1)))
standarddev
```
#2d.
$$se = \frac{s}{\sqrt{n}}$$
```{r}
standarderr <- standarddev/sqrt(length(s1))
standarderr
```
#2e.
$$CI=[\bar{x}-1.96*se,\bar{x}\\+1.96*se]$$
```{r}
lowerl<-mean1-1.96*standarderr
upprl<-mean1+1.96*standarderr
lowerl
upprl
```
#2f.
$$CI=[\bar{x}-1.83*se,\bar{x}\\+1.83*se]$$
```{r}
t<-qt(0.975,9)
t
lowerl1<-mean1-t*standarderr
upprl1<-mean1+t*standarderr
lowerl1
upprl1
```
#2g.
For a larger sample set, the distribution of the sample mean will be normally distributed. As the sample set is small (i.e.n=10) here, there will be more variations within the observations and we cannot say with confidence that the data is statistically significant. 

#3a.
```{r}
n_sample<-c(1:100)
n<-length(n_sample)
m_1<- mean(n_sample)
sd_1<-sd(n_sample)
```

$$CI=\bar{x}\pm1.96*\frac{sd}{\sqrt{n}}$$
```{r}
lowerlimit<-m_1-1.96*sd_1/sqrt(n)
lowerlimit
upperlimit<-m_1+1.96*sd_1/sqrt(n)
upperlimit
```
$$CI'=\bar{x}\pm1.96*\frac{sd}{\sqrt{n'}}$$
Also,
$$1.96*\frac{sd}{\sqrt{n'}}=1.96*\frac{sd}{2* \sqrt{n}}$$
i.e.
$$2*\sqrt{n}=\sqrt{n'}$$
$$n'=4*n=400$$
```{r}
n1 <- 4*n
n1 #Thus, we need 300 more observations from the original n. 
lowerlimit1<-m_1-1.96*sd_1/sqrt(n1)
lowerlimit1
upperlimit1<-m_1+1.96*sd_1/sqrt(n1)
upperlimit1
```
#3b.
$$n=(\frac{1.96*sd}{se})^2$$
```{r}
number1<-round((1.96*20000/1000)^2)
number1
number2<-round((1.96*20000/100)^2)
number2
```

#4.
We pick t distribution because n<=30
```{r}
nruns<-1000
nsamples<-20
sample_summary<-matrix(NA,nruns,3)
for(j in 1:nruns){
  sampler <- rep(NA,nsamples)
  for(i in 1:nsamples){
    if(runif(1) < 0.5){
      sampler[i] <- rnorm(1,40,15)
    }
    else{
      sampler[i] <- rnorm(1,60,27)
    }
  }
  sample_summary[j,1] <- mean(sampler)
  standard_error <- sd(sampler)/sqrt(nsamples) 
  sample_summary[j,2] <- mean(sampler) - qt(0.995,19)*standard_error 
  sample_summary[j,3] <- mean(sampler) + qt(0.995,19)*standard_error 
}
counter = 0
for(j in 1:nruns){
  if(50 > sample_summary[j,2] && 50 < sample_summary[j,3]){
    counter <- counter + 1
  }
}
counter 
```
Thus we can say with approximately 99% coonfidence that the data is statistically significant.
