---
title: "HW05_Sampathirao_A"
author: "Anvita Sampathirao"
date: "6/10/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
#1a. 
$$H_0:\mu\geq2.03$$
$$H_1:\mu<2.03$$

#1b. 
$$t-stat=\frac{\bar{x}-\mu_0}{se}$$
```{r}
serr<-0.07/10
t1<-(1.75-2.03)/serr
t1
```
#1c.
```{r}
lowertail<-qt(0.05,99,lower.tail = TRUE)
lowertail
```
i.e.,
$$RR: (-\infty,-1.66)$$
Because t-statistic of our sample set, -40 lies in the rejection region of t distribution, we can thus reject the null hypothesis.

#1d.
```{r}
lower_lim<-(1.75-(1.64*serr))
lower_lim
upper_lim<-(1.75+(1.64*serr))
upper_lim
```
The 95% confidence interval thus is: $$CI=[1.73852,1.76148]$$
#1e.
```{r}
pt(t1,99,lower.tail=TRUE)
```
$$Since, p-value\leq0.05$$
Thus, null hypothesis can be rejected at 95% significance level

```{r}
#t.test to verify
set.seed(1)
test_sample<-rnorm(100, mean = 1.75, sd = 0.07)
t.test(test_sample,mu=2.03)
```

#2a.
$$H_0:\mu_\text{men}-\mu_\text{women}=0$$
$$H_0:\mu_\text{men}-\mu_\text{women}\neq0$$
#2b.
$$df=2*n-2$$
$$S.E.=s/\sqrt{n}$$
$$t-stat=\frac{\bar{x_1}-\bar{x_2}}{\sqrt{se_1^2+se_2^2}}=\frac{x_1-x_2}{\sqrt{2}*S.E.}$$
```{r}
serr2<-200/sqrt(50)
serr2
alpha<-0.05
t2<-(1124-1245)/(sqrt(2)*serr2)
t2
```
#2c.
```{r}
df<-(2*50)-2
df
v<-c(qt(c(alpha/2,1-(alpha/2)),df))
v
```
i.e.,
$$RR: (-\infty,-1.984467)\cap(1.984467,\infty)$$
Because t-statistic of our sample sets, -3.025 lies in the rejection region of t distribution, we can thus reject the null hypothesis.

#2d.
$$CI_(\mu_1-\mu_2): [(\bar{x_1}-\bar{x_2})\pm1.98*se]$$
$$\bar{x_1}-\bar{x_2}=1124-1245=-121$$
```{r}
lowerlimit<-(-121-(1.98*serr2))
lowerlimit
upperlimit<-(-121+(1.98*serr2))
upperlimit
```
#2e.
```{r}
pt(t2,df)*2
```
$$Since, p-value\leq0.05$$
Thus, null hypothesis can be rejected at 95% significance level

```{r}
#t.test to verify
set.seed(1)
men<- rnorm(50, 1124,200)
women<-rnorm(50,1245,200)
t.test(men,women)
```

#3a.
$$H_0:\mu_t-\mu_c=0$$
$$H_1:\mu_t-\mu_c>0$$
Therefore we use a right tailed t test
```{r}
x_t<-78
x_c<-75
sd_t<-20
sd_c<-5
n<-50
x_diff<-x_t-x_c 
x_diff #mean difference between the 2 groups
se_t<-sd_t/sqrt(n) 
se_t #standard error for treatment group
se_c<-sd_c/sqrt(n) 
se_c #standard error for control group
se_diff<-sqrt((se_c)^2+(se_t)^2)
se_diff #standard error difference between the 2 groups
df1<-(se_diff^4)/((se_t^4/(n-1))+(se_c^4/(n-1)))
df1 #degree of freedom
t3<-x_diff/se_diff
t3 #test statistic
uppertail<-function(alpha1) 
#function to determine if we can reject null hypothesis at the given significance level
{
c<-qt(alpha1,df1,lower.tail = FALSE)
if(t3>=c)
{
  CI<-c(x_diff-(c*se_diff),x_diff+(c*se_diff))
  print("Reject Null hypothesis")
}
else
{
  CI<-c(x_diff-(c*se_diff),x_diff+(c*se_diff))
  print("Fail to Reject Null Hypothesis, because mu_0=0 lies in CI, which is")
  return(CI)
}
}
uppertail(0.10) #significance level 90%
uppertail(0.05) #significance level 95%
uppertail(0.01) #significance level 99%
```
i.e.,
We do not have sufficient evidence with the given data to test our null hypothesis.
```{r}
set.seed(1)
treatment<- rnorm(50, 78,20)
control<-rnorm(50,75,5)
t.test(treatment,control)
```
#4a.
```{r}
library(AER)
data("CPSSW04")
library(ggplot2) 
ggplot(CPSSW04, aes(x = degree, y = earnings)) + geom_boxplot() + theme_classic()
```

#4b.
```{r}
t.test(earnings ~degree, data = CPSSW04)
```
#4c.
```{r}
t.test(CPSSW04$earnings[CPSSW04$age>=25 & CPSSW04$age<=29],
       CPSSW04$earnings[CPSSW04$age>=30 & CPSSW04$age<=34])
```

