---
title: "HW11_Sampathirao_A"
author: "Anvita Sampathirao"
date: "8/5/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressMessages(library(AER))
data(HMDA)
library(ggplot2)
suppressMessages(library(stargazer))
HMDA$deny <- ifelse(HMDA$deny == "yes", 1, 0)
HMDA$afam <- ifelse(HMDA$afam == "yes", 1, 0)
HMDA$phist <- ifelse(HMDA$phist == "yes", 1, 0)
HMDA$selfemp <- ifelse(HMDA$selfemp == "yes", 1, 0)
HMDA$insurance <- ifelse(HMDA$insurance == "yes", 1, 0)
HMDA$condomin <- ifelse(HMDA$condomin == "yes", 1, 0)
HMDA$single <- ifelse(HMDA$single == "yes", 1, 0)
HMDA$hschool <- ifelse(HMDA$hschool == "yes", 1, 0)
HMDA$chist <- as.numeric(HMDA$chist)
HMDA$mhist <- as.numeric(HMDA$mhist)
```

#1

```{r}
plot1<- ggplot(data = HMDA, aes(x= pirat, y= deny)) + 
  geom_point () + 
  geom_smooth(method= "lm", formula = y~x, se= FALSE)
plot1
```

#2

```{r, results='asis', warning=FALSE}
LinReg1<- lm(deny ~ pirat, data = HMDA)
LinReg2<- lm(deny ~ pirat + afam, data = HMDA)
LinRegs<- list (LinReg1, LinReg2)
stargazer(LinRegs, type = "latex", intercept.bottom = FALSE, df = FALSE)
```
In Model (1),
- the estimator for pirat variable is statistically different than 0 and hence, a unit change in payments to income ratio will result in a loan most likely being denied by 60.4%

In Model (2),
- the estimator for pirat variable is statistically different than 0 and a unit change in payments to income ratio will result in a in a loan most likely being denied by 55.9%

- the estimator for afam is also statistically different than 0. A loan is likely to get denied by 17.7% if a person applying for it is an african american as compared to not being an african american

Yes, afam is a relevant omitted variable in Model (1), as we note that the value of beta_pirat has changed from Model (1) to Model (2)

#3

```{r}
PIavg<- mean(HMDA$pirat)
betas<- LinReg2$coefficients
P1.1<- as.numeric(betas[1] + (PIavg*betas[2]) + (1*betas[3])) #When an african american
P1.0<- as.numeric(betas[1] + (PIavg*betas[2]) + (0*betas[3])) #When not an african american
P1.1/P1.0
```

#4

```{r}
#Finding out the order in which variables are to be added by strength of correlation
corData<- cor(HMDA)
corData<- corData[, colnames(corData) %in% c("deny", "afam")]
corData
a<- corData[,1] * corData[,2]
sort(a, decreasing = TRUE)
```
```{r, results='asis', warning=FALSE}
#Running Logit Regressions
LogitReg1 <- glm(deny ~ pirat + afam, 
                 data = HMDA, family = "binomial")
LogitReg2 <- glm(deny ~ pirat + afam + chist, 
                 data = HMDA, family = "binomial")
LogitReg3 <- glm(deny ~ pirat + afam + chist + phist, 
                 data = HMDA, family = "binomial")
LogitReg4 <- glm(deny ~ pirat + afam + chist + phist + insurance, 
                 data = HMDA, family = "binomial")
LogitReg5 <- glm(deny ~ pirat + afam + chist + phist + insurance + lvrat, 
                 data = HMDA, family = "binomial")
LogitReg6 <- glm(deny ~ pirat + afam + chist + phist + insurance + lvrat 
                 + mhist, data = HMDA, family = "binomial")
LogitReg7 <- glm(deny ~ pirat + afam + chist + phist + insurance + lvrat 
                 + mhist + single, data = HMDA, family = "binomial")
LogitReg8 <- glm(deny ~ pirat + afam + chist + phist + insurance + lvrat 
                 + mhist + single + condomin, 
                 data = HMDA, family = "binomial")
LogitReg9 <- glm(deny ~ pirat + afam + chist + phist + insurance + lvrat 
                 + mhist + single + condomin + hirat, 
                 data = HMDA, family = "binomial")
LogitReg10 <- glm(deny ~ pirat + afam + chist + phist + insurance + lvrat 
                  + mhist + single + condomin + hirat + hschool, 
                  data = HMDA, family = "binomial")
LogitReg11 <- glm(deny ~ pirat + afam + chist + phist + insurance + lvrat 
                  + mhist + single + condomin + hirat + hschool + selfemp, 
                  data = HMDA, family = "binomial")
LogitReg12 <- glm(deny ~ pirat + afam + chist + phist + insurance + lvrat 
                  + mhist + single + condomin + hirat + hschool + selfemp + unemp, 
                  data = HMDA, family = "binomial")

LogitRegs1 <- list(LogitReg1, LogitReg2, LogitReg3, LogitReg4, LogitReg5, 
                  LogitReg6)
LogitRegs2 <- list(LogitReg7, LogitReg8, LogitReg9, LogitReg10, 
                  LogitReg11, LogitReg12)

stargazer(LogitRegs1, type = "latex", intercept.bottom = FALSE, df= FALSE)
stargazer(LogitRegs2, type = "latex", intercept.bottom = FALSE, df= FALSE)
```
We see that the value of Beta_afam is changing in every regression until Model (12). Hence, we have to test for multicollinearity.

```{r}
#Check for multi collinearity
aux1_lr12 <- lm(pirat ~ afam + chist + phist + insurance + lvrat + 
                  mhist + single + condomin + hirat + hschool + selfemp + unemp, 
                data = HMDA)
aux2_lr12 <- lm(afam ~ pirat + chist + phist + insurance + lvrat + 
                  mhist + single + condomin + hirat + hschool + selfemp + unemp, 
                data = HMDA)
aux3_lr12 <- lm(chist ~ pirat + afam + phist + insurance + lvrat + 
                  mhist + single + condomin + hirat + hschool + selfemp + unemp, 
                data = HMDA)
aux4_lr12 <- lm(phist ~ pirat + afam + chist + insurance + lvrat + 
                  mhist + single + condomin + hirat + hschool + selfemp + unemp, 
                data = HMDA)
aux5_lr12 <- lm(insurance ~ pirat + afam + chist + phist + lvrat + 
                  mhist + single + condomin + hirat + hschool + selfemp + unemp, 
                data = HMDA)
aux6_lr12 <- lm(lvrat ~ pirat + afam + chist + phist + insurance + 
                  mhist + single + condomin + hirat + hschool + selfemp + unemp, 
                data = HMDA)
aux7_lr12 <- lm(mhist ~ pirat + afam + chist + phist + insurance + 
                  lvrat + single + condomin + hirat + hschool + selfemp + unemp, 
                data = HMDA)
aux8_lr12 <- lm(single ~ pirat + afam + chist + phist + insurance + 
                  lvrat + mhist + condomin + hirat + hschool + selfemp + unemp, 
                data = HMDA)
aux9_lr12<- lm(condomin ~ pirat + afam + chist + phist + insurance + 
                 lvrat + mhist + single + hirat + hschool + selfemp + unemp, 
               data = HMDA)
aux10_lr12<- lm(hirat ~ pirat + afam + chist + phist + insurance + 
                  lvrat + mhist + single + condomin + hschool + selfemp + unemp, 
                data = HMDA)
aux11_lr12<- lm(hschool ~ pirat + afam + chist + phist + insurance + 
                  lvrat + mhist + single + condomin + hirat + selfemp + unemp, 
                data = HMDA)
aux12_lr12<- lm(selfemp ~ pirat + afam + chist + phist + insurance + 
                  lvrat + mhist + single + condomin + hirat + hschool + unemp, 
                data = HMDA)
aux13_lr12 <- lm(unemp ~ pirat + afam + chist + phist + insurance + 
                   lvrat + mhist + single + condomin + hirat + hschool + selfemp, 
                 data = HMDA)
aux1_r2 <- summary(aux1_lr12)$r.squared
aux2_r2 <- summary(aux2_lr12)$r.squared
aux3_r2 <- summary(aux3_lr12)$r.squared
aux4_r2 <- summary(aux4_lr12)$r.squared
aux5_r2 <- summary(aux5_lr12)$r.squared
aux6_r2 <- summary(aux6_lr12)$r.squared
aux7_r2 <- summary(aux7_lr12)$r.squared
aux8_r2 <- summary(aux8_lr12)$r.squared
aux9_r2 <- summary(aux9_lr12)$r.squared
aux10_r2 <- summary(aux10_lr12)$r.squared
aux11_r2 <- summary(aux11_lr12)$r.squared
aux12_r2 <- summary(aux12_lr12)$r.squared
aux13_r2 <- summary(aux13_lr12)$r.squared
vif1 <- 1/ (1 - aux1_r2)
vif2 <- 1/ (1 - aux2_r2)
vif3 <- 1/ (1 - aux3_r2)
vif4 <- 1/ (1 - aux4_r2)
vif5 <- 1/ (1 - aux5_r2)
vif6 <- 1/ (1 - aux6_r2)
vif7 <- 1/ (1 - aux7_r2)
vif8 <- 1/ (1 - aux8_r2)
vif9 <- 1/ (1 - aux9_r2)
vif10 <- 1/ (1 - aux10_r2)
vif11 <- 1/ (1 - aux11_r2)
vif12 <- 1/ (1 - aux12_r2)
vif13 <- 1/ (1 - aux13_r2)
vifs <- list(vif1, vif2, vif3, vif4, vif5, vif6, vif7, 
             vif8, vif9, vif10, vif11, vif12, vif13)
vifs > 10
vifs > 5
```
We can note that for all regressors VIF is less than 5 we can be confident that imperfect multicollienarity is not an issue in regression (12). And, if is not an issue in regression (12) - which includes the larger number of independent variables - then it won’t be an issue in models (1) to (11). Hence, Model (12) is the model with the least bias. 


#5

```{r}
plot2 <- plot1 + geom_smooth(method = "glm", 
                             formula = y ~ x, 
                             method.args = list(family = "binomial"), 
                             color = "red", se = FALSE)
plot2
```
The logistic regression model is a more appropriate specification as it fits the values of deny better than the linear regressor which extends beyond the bounds of deny variable. 

#6

```{r}
betalog <- LogitReg12$coefficients
exp(betalog[3])
```

Because the relation is non-linear, the odds of a loan getting denied for an african american is 2.02 times the odds of a loan getting denied for a non african american.

#7

```{r}
chistavg <- mean(HMDA$chist)
phistavg <- mean(HMDA$phist)
insuranceavg <- mean(HMDA$insurance)
lvratavg <- mean(HMDA$lvrat)
mhistavg <- mean(HMDA$mhist)
singleavg <- mean(HMDA$single)
condominavg<- mean(HMDA$condomin)
hiratavg <- mean(HMDA$hirat)
hschoolavg <- mean(HMDA$hschool)
selfempavg <- mean(HMDA$selfemp)
unempavg <- mean(HMDA$unemp)

invlogit <- function(x){
  yloghat <- betalog[1] + (betalog[2]*PIavg) + (betalog[3]*x) + 
    (betalog[4]*chistavg) + (betalog[5]*phistavg) + (betalog[6]*insuranceavg) + 
    (betalog[7]*lvratavg) + (betalog[8]*mhistavg) + (betalog[9]*singleavg) + 
    (betalog[10]*condominavg) + (betalog[11]*hiratavg) + (betalog[12]*hschoolavg) +
    (betalog[13]*selfempavg) + (betalog[14]*unempavg)
  yhat <- exp(yloghat)/ (1 + exp(yloghat))
  return(yhat)
}
P7.1.1 <- invlogit(1) #When an african american
P7.1.0 <- invlogit(0) #When not an african american
as.numeric(P7.1.1/P7.1.0)

```
The odds of a loan getting denied when an applicant is an african american versus when an applicant is not an african american is 1.89. This is a better estimator than the odds determined with the linear regression model because the range of of the odds ratio is bounded by (0,1)- that fits all values within the domain of deny variable. 
